---
layout: tutorial
title: 'Mini-Hackathon 3: Tidy Scripts'
author: ""
output: pdf_document
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(echo = TRUE)
library(printr)
library(rmarkdown)
```

Mini-Hackathons are performed and submitted in pairs of two. 
You must hand in your assignment on Canvas the next week before **Tuesday Midnight**.

Use this RMarkdown template on the canvas page for this mini-hackathon to complete your hackathon.
When you are finished, knit the file into a pdf with the knit button in the toolbar (or using Ctrl+Shift+K).
For this you need to have the `knitr` and `printr` packages installed, and all your code needs to work (see the R course companion for more instructions).
**If you cannot knit the `.Rmd` file, there is probably an error in your R code, therefore add `eval=FALSE` to the code chunk: `{r, eval = FALSE}`, so you are still able to knit and upload the file.**

# This Mini-Hackathon
This mini-hackathon builds upon the tutorials on [transforming](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/R-tidy-5-transformation.md), [summarizing](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/R-tidy-5b-groupby.md), [reshaping](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/r-tidy-12-reshaping.md), and [visualizing](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/r-tidy-3_7-visualization.md) data. 
For this hackathon, you can (and are strongly recommended to) use code from this week's tutorial as well as provided here.
If you aim to conduct additional analyses in R, we of course encourage this.
Nevertheless, it is important to not do additional analysis just for the sake of running more code chunks. 
For that reason, please provide a justification for these additional analyses.

Also, we recommend to use parameters for RMarkdown codeblocks, in particular the `cache = TRUE` parameter for codeblocks that take long to compute (e.g., downloading data from AmCAT). 
A brief explanation of some usefull parameters to make a `.Rmd` file pretty is given in the [the tutorial of two weeks ago](https://github.com/MarikenvdVelden/Replication-Hackathons/blob/main/Intro-to-rmd-and-data-retrieval.md).
Additionally, you can use [this cheat sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf).

Important to take into account is that this week we will build upon codes of last weeks. 

### Hackathon Challenges
#### Challenge 1
 _Choose a data set._
_Perform the following transformations to the data based on the tidyverse:_

- _Apply a dictionary from `quanteda.dictionaries` to the corpus of your textual data and create a meaningful variable (e.g. [subjectivity score](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/sentiment_analysis.md)) using `mutate`._
- _Subset the data set by (a) filtering out values that are not of interest (e.g. neutral category in sentiment analyses) using `filter`; and (b) selecting the variables that you need for computing new variables (e.g. mean values of a variable), using `select`. _
- _Aggregate (or summarize) the data to another group level, for example to another time level (e.g. monthly level) or based on another variable of your choice (e.g. news outlet) using `group_by`, and `summarize` or `mutate`._



```{r, echo=FALSE, message=F}

```


#### Challenge 2
_Develop an infographic to visualize a meaningful relationship based on your data using `ggplot2`.Try to make the infographic as informative and appealing as possible._

#### Challenge 3
_What does your infographic demonstrate? Explain your answer as if you would inform an Economist reader._

### Challenge 4
_What if you would write for a tabloid newspaper such as the Sun, would you assess, based on he conclusions in the study of [Prior (2014)](https://vu.on.worldcat.org/oclc/8272642139), the visual literacy of the readership to change? If so, what would you adjust in the infographic?_

