---
layout: tutorial
title: 'Mini-Hackathon 3: Tidy Scripts'
author: "Mariken van der Velden & Kasper Welbers"
output:
  github_document:
  toc: yes
editor_options: 
  chunk_output_type: console
---
  
```{r setup, include=FALSE}
## include this at top of your RMarkdown file for pretty output
## make sure to have the printr package installed: install.packages('printr')
knitr::opts_chunk$set(echo = TRUE)
library(printr)
library(rmarkdown)
```

Mini-Hackathons are performed and submitted in pairs of two. 
You must hand in your assignment on Canvas the next week before **Tuesday Midnight**.

Use this RMarkdown template on the canvas page for this mini-hackathon to complete your hackathon.
When you are finished, knit the file into a pdf with the knit button in the toolbar (or using Ctrl+Shift+K).
For this you need to have the `knitr` and `printr` packages installed, and all your code needs to work (see the R course companion for more instructions).
If you cannot knit the `.Rmd` file, there is probably an error in your R code, therefore add `eval=FALSE` to the code chunk: `{r, eval = FALSE}`, so you are still able to knit and upload the file.

# This Mini-Hackathon
This mini-hackathon builds upon the tutorials on [transforming](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/R-tidy-5-transformation.md), [summarizing](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/R-tidy-5b-groupby.md), [reshaping](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/r-tidy-12-reshaping.md), and [visualizing](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/r-tidy-3_7-visualization.md) data. 
For this hackathon, you can (and are strongly recommended to) use code from this week's tutorial as well as provided here.
If you aim to conduct additional analyses in R, we of course encourage this.
Nevertheless, it is important to not do additional analysis just for the sake of running more code chunks. 
For that reason, please provide a justification for these additional analyses.

Also, we recommend to use parameters for RMarkdown codeblocks, in particular the `cache = TRUE` parameter for codeblocks that take long to compute (e.g., downloading data from AmCAT). 
A brief explanation of some usefull parameters to make a `.Rmd` file pretty is given in the [the tutorial of two weeks ago](https://github.com/MarikenvdVelden/Replication-Hackathons/blob/main/Intro-to-rmd-and-data-retrieval.md).
Additionally, you can use [this cheat sheet](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf).

Important to take into account is that this week we will build upon codes of last weeks. 

### Hackathon Challenges

#### Challenge 1
Choose a data set. 
You can either work with the AmCAT data set with newspaper articles on the US presidential candidates in 2020 or the Guardian data set with newspaper articles in politics, economy, society and international sections from 2012 to 2016.
To get the data, one can run the code to download the data from `quanteda.corpora` or download the `.csv` file with AmCAT data from Canvas.
Take a look at respectively [the first tutorial](https://github.com/MarikenvdVelden/Replication-Hackathons/blob/main/Intro-to-rmd-and-data-retrieval.md) or the first mini-hackathon (see [here](https://github.com/MarikenvdVelden/Replication-Hackathons/blob/main/Mini-Hackathon1.md) and [here](https://github.com/MarikenvdVelden/Replication-Hackathons/blob/main/Mini-Hackathon1-Guardian.md)) to see how to import data.
Use tidyverse to transform the data in the following ways:

- Apply a dictionary from `quanteda.dictionaries` to the corpus of your textual data and create a meaningful variable (e.g. [subjectivity score](https://github.com/ccs-amsterdam/r-course-material/blob/master/tutorials/sentiment_analysis.md)) using `mutate`.
- Subset the data set by (a) filtering out values that are not of interest (e.g. neutral category in sentiment analyses) using `filter`; and (b) selecting the variables that you need for computing new variables (e.g. mean values of a variable), using `select`. 
- Aggregate (or summarize) the data to another group level, for example to another time level (e.g. monthly level) or based on another variable of your choice (e.g. news outlet) using `group_by`, and `summarize` or `mutate`. 

If you want to use the following code to query the AmCAT data set based on the `.csv` file:
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(amcatr)
library(quanteda)

d <- read_csv("../amcat_data.csv")
corp <- corpus(d, docid_field = 'id', text_field = 'text')
q  <- grepl('econom', texts(corp), ignore.case = T)
d <- corpus_subset(corp, q) %>%
  convert(corp, to = "data.frame")
```

Querying the Guardian data is described in [Mini-Hackathon 2](https://github.com/MarikenvdVelden/Replication-Hackathons/blob/main/Mini-Hackathon2.md).

#### Challenge 2
Develop an infographic to visualize a meaningful relationship based on your data using `ggplot2`.
Try to make the infographic as informative and appealing as possible.
For this, you can use the [_ggplot2 cheat sheet_](https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf), the [_R Graphics Cookbook_](http://www.cookbook-r.com/Graphs/), and/ or the [_BBC Visual and Data Journalism Cookbook_](https://bbc.github.io/rcookbook/).
If you want to make more than one graph, have a look at the `facet_wrap` function: `?facet_wrap`.

#### Challenge 3
What does your infographic demonstrate? Explain your answer as if you would inform an Economist reader.

#### Challenge 4
What if you would write for a tabloid newspaper such as the Sun, would you assess, based on the conclusions in the study of [Prior (2014)](https://vu.on.worldcat.org/oclc/8272642139), the visual literacy of the readership to change? If so, what would you adjust in the infographic?



